### 2025-04-08 main.tex
changes from john

%% 
%% Copyright 2007-2024 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{lineno}
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\journal{Journal of Computer Languages}

\begin{document}

\begin{frontmatter}

 \author{Shuai Fu}

 \author{Tim Dwyer}

 \author{Peter J. Stuckey}

 \author{John Grundy}

 \affiliation{organization={Monash University}, country={Australia}}

\title{GeckoGraph: A Visual Language for Polymorphic Types}

%% Abstract
\begin{abstract}
Polymorphic types are an important feature in most strongly typed programming languages. They allow functions to be written in a way that can be used with different data types, while still enforcing the relationship and constraints between the values. However, programmers often find polymorphic types difficult to use and understand and tend to reason using concrete types. We propose GeckoGraph, a graphical notation for types. GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier. We conducted a large-scale human study using GeckoGraph compared to text-based type notation. To our knowledge, this is the largest controlled user study on functional programming ever conducted. The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.  
\end{abstract}

%% Keywords
\begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
Polymorphic Types \sep Functional Programming \sep Visual Languages \sep Visualization
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction} \label{sec:intro}
In programming languages, a polymorphic type \cite{Cardelli1987-fp} can represent values of different types while providing an interface to describe common behaviors for those values. Polymorphic types are central to the succinctness of contemporary statically-typed languages, enabling a considerable degree of type-safe abstraction and reusability. Parametric types are available in many programming languages, from functional languages such as Haskell and ML to imperative and multi-paradigm languages such as Rust\cite{Klabnik_undated-wx} and Go\cite{Griesemer_undated-ff}. Polymorphism allows programs to be written in a way that is more generic and adaptable to different data types, enabling greater flexibility and code reuse. Polymorphic types are ideal for modeling abstractions, such as properties of mathematical objects and laws that hold on these objects.

While promoting robustness and code reusability, polymorphic types present challenges in comprehension, particularly for novice users \cite{Jun2000-ec, Jun2000-yu}. Research indicates that humans tend to focus on concrete types and rely on polymorphic type checking only as a last resort. Moreover, polymorphic types often complicate type error usability. During type checking, compilers frequently generate new polymorphic type variables temporarily. These variables are typically discarded after type checking unless a type error occurs, in which case they are referenced in error messages. Consequently, programmers often face the task of resolving type errors involving type variables they did not explicitly author (Fig. \ref{fig:example-foldable}).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Foldable}
  \caption{\label{fig:example-foldable} An example type error where a programmer mistakenly provided a \texttt{Char} instead of a \texttt{String} literal. \textbf{Left --} The compiler shows an error message comparing the provided Char type to a confusingly named type \texttt{t0 a0}. \textbf{Right --} GeckoGraph shows the exact message with the two types in graphic notation, highlighting the structural difference rather than identifier names.}
\end{figure}

Polymorphic type is one of the oldest topics in programming language theory \cite{Cardelli1987-fp}. However, little research focuses on the \textit{usability} of polymorphic types. It is often argued that the expressiveness and power of type systems often come at the cost of usability~\cite{Hage2020-hg}. We aim to investigate the challenges of using polymorphic types and explore how to improve their usability with visualization and modern HCI techniques. In this paper, we propose GeckoGraph, a graphical notation for types. GeckoGraph aims to complement traditional text-based type notation and make reading, understanding, and comparing types easier. GeckoGraph is prototyped and verified iteratively, leading to a design with visual clarity applicable to many programming contexts. Our study evaluating 714 participants' ability to solve a series of programming challenges using GeckoGraph versus text-based type annotation is, to our knowledge, the largest controlled user study of a functional programming tool or concept. The study results show that GeckoGraph helps improve programmers' ability to succeed in resolving the programming challenges, especially for novice programmers.

\section{GeckoGraph}

GeckoGraph is a visual notation for type annotations in statically typed programming languages. It is intended to work tangibly with text-based annotations but uses colors, shapes, and symbols to make structures of types easy to identify at a glance. In this section, we describe the design of GeckoGraph and highlight some unique benefits of programming with GeckoGraph.

\subsection{Design of GeckoGraph}
The design of GeckoGraph focuses on visualizing types in functional languages (e.g., Haskell, ML). In this paper, we use Haskell as an example. As illustrated in this section, it can express basic types, polymorphic types, algebraic data types, and many advanced type-level features. However, GeckoGraph could also be used in imperative and multiparadigm languages such as TypeScript and Rust. We provide a GeckoGraph visualization tool avaiable online for multiple languages \cite{Fu2025-ka}.

We identified three main design goals for GeckoGraph based on the challenges of using polymorphic types~\cite{Jun2000-ec, Jun2000-yu} and how programmers tend to use type annotations to support their programming tasks~\cite{Justin_Lubin2021-yy}, as follows.

\paragraph{\textbf{(D1) Easy to learn}}\label{goal1} GeckoGraph should take little to no effort to learn. The rules to translate a text-based type notation to GeckoGraph should be minimal. Where possible, GeckoGraph should be intuitive to programmers who are familiar with text-based type notation.

\begin{figure}[!h]
  \includegraphics[width=\linewidth]{figures/Design}
  \caption{
        \label{fig:design}
        Examples of various types as represented in GeckoGraph, including type variables (A) and concrete types (B). Data types (C, D, E), function types (F, G, H), and type classes (I, J).
  }
\end{figure}

\paragraph{\textbf{(D2) Easy to parse for humans}}  \label{goal2} GeckoGraph should make the task of reading and understanding type notation easy. It should emphasize the less obvious properties of a type signature. GeckoGraph should eliminate the need for mental backtracking, such as counting opening and closing parentheses and remembering which type classes are required on which variables.

\paragraph{\textbf{(D3) Optimized for comparing and searching}} \label{goal3} GeckoGraph aims to make the task of comparing two types easy, especially to make subtle differences in text-based notation harder to miss. This also includes the task of choosing an ideal function from a list of potential functions. For example, programmers search for a desired function from a documentation site with only partial knowledge of its type (e.g., the arity of a function, containing a known type, or type class requirements).

\paragraph{Simple Types} 
Simple types, such as type variables and concrete types, are displayed in a cell \includegraphics[height=1em]{figures/SimpleType}: a solid-colored rectangular box with an angled corner on the top left. Each type identifier encodes a distinct color hue of the cell. Its first 1 or 2 letters are displayed inside the cell at the bottom left to provide familiarity (design goal \ref{goal1}) and a strong secondary encoding. The angled corner in the top left provides visual separation between two cells, even when the same color cells are next to each other, allowing GeckoGraph to be zoomed out to extremely small sizes (Section \ref{subsec:space}) without suffering readability (design goal \ref{goal2}).

\paragraph{Complex types}
GeckoGraph displays an algebraic data type as a larger cell, where the type constructor half encloses its arguments: \includegraphics[height=1em]{figures/DataType.png}. The arguments are aligned in the bottom right of the cell. Two distinct visual dimensions are used to provide additional visual clarity (design goal \ref{goal2}). Data types containing more arguments (e.g., \texttt{ a b c} or \texttt{(a b) c}) will expand horizontally \includegraphics[height=1em]{figures/DataTypeWide.png}. Data types that are nested (e.g. \texttt{ a (b c)}) will grow taller \includegraphics[height=1.2em]{figures/DataTypeNested.png}. This distinction accommodates our design goal \ref{goal3}. Note that the height of GeckoGraph grows only upwards, but not downwards. Not only does this allow GeckoGraph to be more efficient in its space usage, but it also allows the legend text to be consistently aligned at the bottom and can be read as though it is a text-based type notation (design goal \ref{goal1}).

\paragraph{Function Types}
Functions are the fundamental building blocks of functional programming languages, and function types are ubiquitous and the most important in type-level programming. In Haskell, \texttt{(->)} is defined as an infix type operator with the right associativity to provide succinct type annotation. GeckoGraph preserves this syntax feature to make the notation more intuitive (design goal \ref{goal1}): the 2 arguments of a function type in GeckoGraph are placed on both sides of the cell \includegraphics[height=1em]{figures/Function.png}. A special function indicator (\texttt{$>>>$}) is displayed at the top of the cell.

Curried functions (e.g., \texttt{ a -> b -> c}) are displayed as two cells of functions merged together \includegraphics[height=1em]{figures/Curry.png}. It appears as if the second function covers the right half of the first function, indicating that the second function is the return type of the first. Regular high-order functions (e.g., \texttt{(a -> b) -> c} ) follow the rules of functions and nested data types \includegraphics[height=1.2em]{figures/HOF.png}. The placement of function indicators aims to make it easy to find desired functions in the documentation site based on function arity and high-order functions (design goal \ref{goal3}). It is easy to tell high-order functions from the vertical position of its function indicator. Similarly, it is easy to count the arity of a function by counting the number of horizontally connected function indicators (Fig. \ref{fig:indicator}).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Indicator}
  \caption{
        \label{fig:indicator}
        An example of visual hierarchy of the function indicator. The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators. For higher-order functions where functions are arguments or return values of other functions, it is very easy to see the ``level" of functions. 
  }
\end{figure}

\paragraph{Type Classes} 
Type classes are an intrinsic part of Haskell \cite{Hudak2007-kn}, and many other functional languages. In GeckoGraph, the type classes (e.g., \texttt{ (A a, B a) => a}) are indicated in the extended area below one or more GeckoGraph cells \includegraphics[height=1.2em]{figures/TypeClass.png}. Each type class required on a type variable displays as a square indicator aligned on the right of the extended area. In GeckoGraph, a type-class gliph is embedded in every type variable that requires this class. This means when displaying the type \texttt{(==) :: Eq a => 
a -> a -> Bool} in GeckoGraph, the constraint \texttt{Eq} appears in twice on both occurrences of the type variable \texttt{a}.

The GeckoGraph type class's design promotes the type class placements rather than the type class names. Programmers can easily see where and how many type classes are required, but they may need an extra step (global legends of the color mapping or pop-up window) to identify the name of the type class. We believe that this is a well-justified trade-off. The strength of this approach lies in situations when type class context is complex. When reading a type \texttt{(A a, A c, B a, B b, C b) => a -> b -> c}, programmers may need to switch back and forth to remember which type classes are needed on which variable. GeckoGraph uses approximity to help minimize the effort to associate each type variable with all its type classes (design goal \ref{goal2}).

\subsection{Key Design Criteria of GeckoGraph}\label{sec:benefits}

GeckoGraph was designed through many different iterations. Many research methods were used to verify ideas, including prototyping, cognitive walk-throughs, and formative studies. We discuss some notable visual dimensions in GeckoGraph and their justifications.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/PreviousVersions}
  \caption{\label{fig:previous}Previous Versions of GeckoGraph. Different encodings represent named types, type variables, type constructors, and high-order functions.}
\end{figure}

\paragraph{Visual Dimensions of GeckoGraph}
\textbf{The colors of GeckoGraph} help programmers to see the permutation of type variables in the input and output of a function. A recent review \cite{Zeng2023-jz} of 59 graphical perception articles showed that combining solid color hue in a filled shape provides stronger visual perception for nominal data such as type identifiers. One example of GeckoGraph's use of color is the ``rotation" function  (Fig. \ref{fig:rotate}) in the user study (Section \ref{sec:evaluation}). With text-based type notation, programmers often rely on mnemonic devices such as alphabetic ordering or naming conventions. For example, the rotation function \texttt{f2 :: Zero a b c d -> Zero b c d a} in the game is less recognizable if changed to \texttt{f2 :: Zero e v m h -> Zero v m h e}.

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/rotate}
  \caption{\label{fig:rotate} The `rotate' function in level 8 of the user study. The name given in the game is `f2'. It shuffles the type arguments of a Zero type}
\end{figure}

\textbf{The horizontal axis of GeckoGraph} often are helpful in identifying differences in function arities. For example, in Fig. \ref{fig:add3}, the programmer intended to implement a function that sums 3 integers. In the implementation, the programmer missed a \texttt{(+)} function at the end; the resulting function type is largely different in length. It is also clear that the function needs to apply to one more binary function to satisfy the length requirement.

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/Add3}
  \caption{\label{fig:add3} An implementation of function \texttt{add3} but the author missed an (+) from the correct implementation (.) ((+) .) (+). GeckGraph highlights the difference in arity and reveals that a binary function is needed on the right-hand side for the arity to match. }
\end{figure}

\textbf{The vertical axis of GeckoGraph} often helps identify the number of levels in a nested complex type. This can often be very useful when inspecting mismatching type errors where data types are nested in different levels. Common examples include when programmers forget to apply the value to ``return" in a monadic block or to use \texttt{liftIO} to cast an \texttt{IO} effect. In Fig. \ref{fig:maybe}, the uses of \texttt{return} are excessive. It can be easily identified by examining the difference in the vertical layers of the two types. In text-based type notation, this is distinguished by different pairs of parentheses. However, parenthesis is an overloaded syntax in type notation. In Haskell, parentheses are used to enclose tuples \texttt{(a, b)}, specify the fixity \texttt{ (a -> b) -> c}, or have no effect \texttt{a -> (b -> c)}.

\begin{figure}[]
  \includegraphics[width=0.6\linewidth]{figures/Maybe}
  \caption{\label{fig:maybe} The function \texttt{f} is planned to have the type \texttt{Maybe a -> a -> Maybe a}. The programmer mistakenly applied the result to the \texttt{return} function, making the result inside a Monad instance.  GeckoGraph reveals the difference in the ``layers" of types. }
\end{figure}

\paragraph{Support Advanced Type Features}
The design of GeckoGraph enables the visualization of many advanced type-level features. \textbf{Kind visualization}: Kind system \cite{Jones1995-lm} is used in Haskell to indicate whether a type is type constructor (e.g. [], Maybe) or a type constant (e.g. Bool, Char), because of this, kind is often described as the type of types. GeckoGraph infers the {\it kind} of type variables and consistently displays this information in GeckoGraph. For example, in figure \ref{fig:advanced} (A), the variable \texttt{a} is inferred to have the kind \texttt{* -> *} because of its use on the right-hand side. GeckoGraph respects this {\it kind} information and displays the same kind level on the first occurrence of \texttt{a}, but over an empty structure, indicated using a dotted outline. \textbf{Qualified constraints}: GeckoGraph's type class notation naturally extends to support qualified constraints. In the type \texttt{forall b. A (a b) => a b}, GeckoGraph shows the scope type class requirement on \texttt{a b} (Fig. \ref{fig:advanced} B).
\textbf{Multiple Parameter Type Class}:  GeckoGraph supports multiple parameter type classes by using multiple shapes with the same color hue to indicate the different parameters of the same type class. For example, for the type \texttt{A a b => a b},  GeckoGraph shows that the variables a and b both need an A class, but they are the different parameters of A (Fig. \ref{fig:advanced} C).

\paragraph{Provide Precise Interactivity}
Modern programming environments often allow programmers to mouse over part of the source code to query detailed information, such as definition, references, or documentation. However, with text-based source code, it is often hard to distinguish whether programmers want the most specific fragment under the cursor or larger blocks. Because of its graphical layout, GeckoGraph allows programmers to precisely select which part of a type signature they intend to query, that is, in Fig. \ref{fig:advanced} (D) when the user mouses over the type class box (orange square) under the second occurrence of \texttt{a} the type class it represents is revealed in detail, a sepecific query that is hard to hard to execute with point and click over text based type annotation.

\begin{figure}[h]
\label{fig:advanced}
  \includegraphics[width=\linewidth]{figures/Advanced}
  \caption{
  \label{fig:advanced}
  Advanced features of GeckoGraph. (A) GeckoGraph supports Kind Visualization if the inferred kind is greater than \texttt{*}. (B) GeckoGraph supports qualified constraints by extending the extended area across multiple type variables. (C) GeckoGraph supports Multiple Parameter Type Classes, using different shapes of the same color to indicate that multiple variables must satisfy certain type classes collectively. (D) GeckoGraph supports the precise selection of its sub-structures. }
\end{figure}

\section{Evaluation}
To evaluate the usefulness of GeckoGraph, we designed and conducted a controlled experiment in the form of a game called "Zero to Hero". The game contains 10 levels of varying difficulty. At each level, participants are asked to implement a function named "\texttt{zeroToHero}" with the help of a limited list of provided functions. These functions are different at each level, and the target types of Zero and Hero vary at each level as well. The details of each level are provided in the Appendix (Appendix \ref{levels}).

The experiment aims to study how programmers use and reason about polymorphic type when performing programming tasks. In particular, we studied how programmers scan and select potentially useful functions from a library and compare intended types and actual types during type errors.

\subsection{An example level}
We illustrate the task of the user study using level 4 of the game. At this level (Fig. \ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a b -> Hero b b}. The available functions are \texttt{f1::Zero a b -> Hero b a}, \texttt{f2::Zero a a -> Hero a a}, \texttt{f3::Zero a b -> Hero b a}, and \texttt{f4::Zero a b -> Hero b b}. Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions. The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).

\begin{figure}[h]
  \includegraphics[width=\linewidth]{figures/GamePlay}
  \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero. On this level -- level 4 (Shown at F) -- the players need to implement the function \texttt{ zeroToHero :: Zero a b -> Hero b b} (B). They write their own definitions in the code editor (D) using a set of provided functions (G). The inferred type of their current definition is shown in (C). When ready, they can test their solution by clicking on the \textit{Attempt} button (A). They can also skip a level by clicking on the \textit{Bypass} button next to it. The output from the compiler, if there is any, is shown in a window below (E). The GeckoGraph in the screenshot uses a different color scheme that is optimized for computer screens.}
\end{figure}

At each level of the game, the programmer must select the right set of functions and compose them correctly to achieve the target result. Not all provided functions are useful. In particular, for this level, participants must discover that only \texttt{f4} and \texttt{f2} are necessary to produce the desired results. One implementation that satisfies the target type is \texttt{zeroTohero z = f2 (f4  z)}.

\subsection{Recruitment}
Participants were recruited online through the Haskell community on Reddit and Discord. Participation is fully anonymized; detailed ethical implications of these experiments were reviewed and approved by the IRB of the authors' institution.

\subsection{Group Assignments}

The experiment uses a between-subject design. Participants are assigned to one of two groups. The tasks assigned to the two groups are in the same order. Participants receive both treatments (with and without GeckoGraph) during their runs but in a different order.   Group one is assisted by GeckoGraph on even levels and text-only type annotation on odd levels. Group two participants are the same but with the order flipped. The number of participants in the two groups is counterbalanced.

\subsection{Hypothesis}
In programming tasks that involve reading and understanding polymorphic types, graphic notation using visual elements that provide higher grouping strength (colors, shapes, sizes, and symbols) can improve the performance of such tasks compared to traditional text-based type notation. Our null hypothesis is: {\it Using graphic notation has no effect compared to traditional text-based type notation.} This hypothesis and the task design were registered at the Open Science Foundation prior to data collection \cite{Stuckey2023-vl}.

\subsection{Task Design} \label{subsection:task}
During the study, participants receive 10 tasks, represented by the 10 levels of the game {\it zeroToHero}. The tasks start off easy but gradually increase in difficulty.  In each task, a target type of the function \texttt{zeroToHero} is shown to the participants. Participants are provided with a list of available functions to implement the target function and to ensure it assumes the target type. This is to simulate the tasks of selecting useful functions from a library. Participants are not allowed to use any other functions or variables outside the provided functions; even the Haskell prelude is not available. This ensures that everyone has the same knowledge and minimizes the effect of familiarity.

Participants can skip a level during the game if they are stuck. We believe that it is normal for anyone to get stuck on a challenging task, and being stuck on one of the 10 tasks does not discount their feedback of the tool. We limit the number of skips that a participant can use during the game to four so that submitting qualitative feedback without completing at least some levels is impossible.

\subsection{Measurements}
During the study, the time spent by participants on each task is recorded. We also record the resulting status of each level, whether it is a success or failure. Before each run, participants nominate their level of Haskell experience on a four-level scale: beginner, familiar,  knowledgeable, and expert.  If a participant has completed all 10 levels (with the help of skipping), we invite the participant to complete a post-study survey. In it, we ask for their opinion on how intuitive the GeckoGraph design is, how distracting they find GeckoGraph, and how helpful GeckoGraph is during the game, using a seven-point scale. In the end, we ask a few open-ended questions, inviting participants to provide their experience using GeckoGraph and their expectations about the potential applications of GeckoGraph.

Data collection from the study was stopped after the planned cut-off period of 14 days. After the cut-off date, the ZeroToHero game is open source and available for free evaluation \cite{Anonymous_undated-ne} and repeating our experiment, but no further data was collected.

\section{Results}

During the data collection period, a total of 714 users participated in the study. Among them, 245 are novice users, 216 are familiar with Haskell, 216 are knowledgeable users, and 88 are expert users.

\subsection{Time to complete levels}

The 10 levels are designed to increase difficulty gradually. From the results of the experiment, most of the tasks align with this trend. However, three tasks stand out in Fig. \ref{fig:level-time}.  Level 7 (mean = 334 seconds) is the hardest task in the game in terms of time, followed by level 8 (mean = 228 seconds) and level 5 (mean = 224 seconds). To complete an average level, the beginner group uses an average of 100 seconds, the familiar group uses 90 seconds, the knowledgeable group uses 80 seconds, and the expert group uses 70 seconds. This roughly aligns with self-reported expertise. We show that the task time on each level follows normal distributions using a Shapiro-Wilk test \cite{Shaphiro1965-dx} (p-value  $ \leq 1.018 \times 10^-16$, for an alpha value of 0.05, p less than 0.05 is considered normal distribution).

Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience. Level 3 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple. Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class. The authors speculate that the more experienced participants are much more familiar with these functions, hence the strong contrast on these three levels.

However, when comparing the task time between the two treatments, we were unable to reject the null hypothesis. In a two-sample T-test, we could not find any significant difference between the two groups overall (p-value = 0.457), nor does there differ between the two groups in any of the four levels of experience (beginner: p-value = 0.845, familiar: p-value = 0.524, Knowledgeable: p-value = 0.712, expert p-value = 0.771).

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/LevelTime}
  \caption{\label{fig:level-time} Time spent on each level, with 95\% confidence interval. We show that the difficulty steadily increases across the game, but levels 5, 7, and 8 are significantly harder than the authors intended. The overall task time of each group roughly matches experience level.}
\end{figure}

\subsection{Success rate}
We saw that, overall, GeckoGraph provides a higher success rate (96.88\%) than text-based type notation (94.62\%). This trend can be seen in every experienced group: beginner group (95. 12\% vs. 92. 68\%), familiar group (97. 39\% vs. 93. 34\%), knowledgeable group (96. 82\% vs. 96. 06\%), and expert group (98. 2\% vs. 96. 40\%). We saw the significance decrease as the user's experience increased. When performing a proportion test on each group, we see that the effect is most significant with the beginner group and reject the null hypothesis (z score = 2.0228, p-value = 0.0431), followed by the familiar group (z score = 1.7495, p-value = 0.0802). The knowledgeable group (z score = 1.0295, p-value = 0.3032) and the expert group (z score = 0.8660, p-value = 0.3756) show less significant differences between treatments.

When breaking down the result in each task (Fig. \ref{fig:success-rate}), we were able to reject the null hypothesis in task 10 of the beginner group and task 10 of the familiar group \ref{fig:success-rate}. We will address this correlation in Section~\ref{sec:discussion}.

Level 10 is the most cognitively demanding: the provided functions are the lengthiest, and the solution is challenging and deceptive. Programmers must use the provided rotate function {\it f2} to adjust the positions of the four components of the {\it Zero} value four times to get to the correct position. Completing the level requires applying the functions {\it f1}, {\it f2}, and {\it f3} in a nested sequence that is 8 layers deep. Consequently, programmers need to frequently switch contexts to recall the purpose of each function. We speculate that this level demands the mental flexibility typical of a Haskell expert to internalize these functions quickly. GeckoGraph helps less experienced programmers to achieve the same level of comprehension fluency by adding structural clarity.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/SuccessfulRate}
  \caption{\label{fig:success-rate}{\bf Success rate of each task with and without GeckoGraph, grouped by experience level.} The figure is cropped from 70\% to 100\% for readability. In most tasks, GeckoGraph provides a small edge. However, significant differences were found in task 10 of the beginner group and task 10 of the familiar group. }
\end{figure}

\subsection{Qualitative Feedback}
In their responses to the post-study survey, most programmers believe that the design of GeckoGraph is intuitive and that its appearance in the interface does not cause distraction.
For the question ``Do you find the GeckoGraph distracting", most of the participants rated a negative score, with an average of 2.88 (Fig. \ref{fig:qualitative} left). For the question ``How intuitive do you find the GeckoGraph?", we saw a reverse correlation of experience (Fig. \ref{fig:qualitative} Middle): experts find the GeckoGraph most intuitive (5.07), followed by the knowledgeable group (4.87), and the familiar group (4.80). The beginner group found it to be the least intuitive but still rated a positive score of 4.71. 
When answering the question ``How helpful do you find GeckoGraph in finding the solution during the game?", the answer is more divided into different experience groups (Fig. \ref{fig:qualitative} right). It is slightly positive for beginners (4.25) and slightly negative for the other groups, the familiar group (3.86) and the knowledgeable group (3.32). The expert group finds that GeckoGraph is relatively unhelpful, with an average score of 2.95.

Two interesting trends emerge. First, the more experienced the users are, the more likely they are to find GeckoGraph intuitive and easy to use. This aligns with our design decision to make GeckoGraph resemble text-based type annotation. Those familiar with text-based type annotation find GeckoGraph second nature to read. Second, the correlation reverses when it comes to perceived helpfulness. The more experienced the users are, the less likely they are to find GeckoGraph helpful. We speculate this pattern occurs because users who are experts in text-based type annotations naturally prefer reading text-based notation and tend to ignore GeckoGraph, diminishing its usefulness. Conversely, novice users are more likely to be drawn to GeckoGraph and find that its visualization offers more clarity and useful insights.

% To quote some feedback from participants: ``Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature. " ``It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is."

% To quote a participant, ``GeckoGraph is quite intuitive to see the permutations of the arguments. Also, to see how to produce and consume arguments."

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Qualitative}
  \caption{\label{fig:qualitative} {\bf The users rated scores of how intuitive (left), distracting (middle), and helpful (right) they found GeckoGraph.} Overall, programmers consider GeckoGraph to be intuitive and not distracting. However, opinions are split on its helpfulness. }
\end{figure}

\subsection{Threats to validity}

\paragraph{Task design}
In our human study, most of the provided functions are very abstract. These functions are created by the authors solely for the gamified study. They are designed to be different from well-known Haskell functions to minimize the familiarity variable. They are also designed with an interest in being challenging and fun. These functions may not be the most representative of real-world Haskell programming.

\paragraph{The use of skips}
Although we justified the use of skips in Section \ref{subsection:task}, the availability of skipping does allow users to adopt more utilitarian strategies, often involving skipping a level without giving it a fair try. This happened more often in the later levels when users realized they had enough skip opportunities left to ``complete the game". These strategies may result in lower recorded success rates than if no skips were allowed.

\section{Discussion} \label{sec:discussion}
\subsection{Strengths}
From the results of our experiment, we see that using GeckoGraph has a significant effect on the success rate of our participants, especially on less experienced programmers. We also see that with the data we collected, we did not find a significant time difference between programming with and without GeckoGraph. To extrapolate the observed expressiveness, we speculate on the practical benefits of programming with GeckoGraph.

\subsubsection{Identify the Most Important Features}
One trend we saw from the qualitative feedback is that programmers find GeckoGraph helpful for finding patterns and important features of the types. Programmers are very positive about GeckoGraph's ability to reveal the most helpful features of a type in distinctive visual elements such as color, length, and height.
To quote some feedback from participants: ``{\it Types are much easier to understand by the GeckoGraph than by trying to parse all parentheses and understand the types from the signature.} " ``{\it It makes it easier to see at a glance when your output type is correct or what the difference between the current type and the target is.}"

\subsubsection{Generalize patterns in type classes}
A common source of confusion among novice Haskell programmers is distinguishing between functions that operate on generalized types and those that work with concrete types \cite{Tirronen2015-nr}. For instance, the type constraints {\tt Integral a => a} and {\tt Integer}, {\tt Floating a => a} and {\tt Float}, as well as {\tt Foldable} instances versus {\tt []} can be particularly challenging. Many Haskell learners may find the differences between concrete types and their generic counterparts irrelevant to their programming tasks, thereby complicating their quest for solutions. In the example shown in Fig. \ref{fig:example-foldable}, a programmer mistakenly supplies a \texttt{Char} instead of a \texttt{String} literal. The compiler responds with an error message that compares the provided \texttt{Char} type to a confusingly named type \texttt{t0 a0}. In this case, the information has been generalized too much and the programmer can hardly find their bearing to the original problem. This level of generalization can undermine a programmer's confidence in understanding the language and navigating type errors. In GeckoGraph, both list types and types with a Foldable instance share a similar structure. Therefore, the type \texttt{t0 a0} in the error message matches the shape of \texttt{[a]}. Bring the programmer closer to the real issue, trying to match a complex type with a concrete simple type. This feature lowers the entry barrier for languages like Haskell, where even basic arithmetic functions {\tt +} and {\tt -} are defined on generic types. This approach lowers the barrier to introducing foundational language concepts without sacrificing generality.

\subsubsection{Consistent color scheme}\label{par:color-scheme}
A common task in programming is to scan for a desired function from a sea of potentially useful functions, such as library documentation. During scanning, programmers often have partial knowledge of the desired function, e.g., the arity, one of the argument types, or the type class it must fulfill.  A typical example is conversion: using a known \texttt{String} type to produce a desired \texttt{Data.Text}  type. Another example is the `lookup' function: using a known \texttt{Data.Map a b} to produce a desired \texttt{b} type. GeckoGraph supports this task by using consistent colors for the same type identifier. Programmers can rely on the color grouping to scan for the desired type in their project or in third-party library documentation.

\subsubsection{Low barrier to learn and understand}
GeckoGraph has some key similarities to traditional text-based type notation. GeckoGraph respects the left-to-right reading order. GeckoGraph uses the familiar symbolic name as the secondary encoding. GeckoGraph simulates the prefix notation in type constructors and the infix notation in type operators. With these considerations, we ensured that programmers familiar with the Haskell language had no problem understanding and using GeckoGraph fluently with minimal training.

This is reflected in our user study feedback. Many participants mentioned that they did not have any prior knowledge of Haskell but were able to solve the puzzles with the help of GeckoGraph.
``It is similar enough to traditional types that it is intuitive." ``This was how I parse the textual representation of types" was pointed out by multiple participants.

\subsubsection{Language Agnostic}
GeckoGraph can be implemented in any language that exercises static typing. We identified applications of this feature in programming and education. Programming projects often use more than one language. Situations include projects using foreign function interfaces or multiple languages to accommodate a client-server architecture. In these cases, GeckoGraph provides a shared notation to describe the functionality and features of programming constructs despite their different text-based notations. In teaching and learning programming languages, GeckoGraph removes the nomenclature difference in different programming languages. For example, when describing the product type in algebraic data types, different language communities use various names: record, named tuple, struct, etc. GechoGraph helps to unify the same programming concept and ignore the minute nomenclature differences.

\subsection{Weaknesses}
\subsubsection{Space Usage}\label{subsec:space}
GeckoGraph uses horizontal space in proportion to the size of the type signature syntax tree, and GeckoGraph uses vertical space in proportion to the depth of the syntax tree. Compared to the traditional text-based language, GeckoGraph has the limitation of requiring vertical space. We have identified some approaches to minimize space usage while retaining most of the advantages of using GeckoGraph, such as displaying only the color blocks without the secondary encoding of identifier names.

\subsubsection{Color Encoding}
GeckoGraph relies highly on color as the primary encoding. It provides a strong visual grouping  \cite{Zeng2023-jz}  for programmers to identify patterns in types, such as the order and placement of substructures. However, the perception of color is different from person to person. This becomes an even bigger issue for visually impaired programmers. Although GeckoGraph uses color-blind friendly schemes, it is only a method to avoid indistinguishable types and is not a strong guarantee of effectiveness. For this, we are exploring different visual encoding methods, such as patterns and shapes, to maximize the accessibility of GeckoGraph.

\subsection{Gamified Human Study}

It is important to recognize that the human study is designed to be a series of puzzles. The tasks are meant to contain entertaining values. We practiced multiple gamification techniques: levels, story/theme, and goals/rewards. \cite{Hamari2014-mc} This not only allowed us to have confidence that participants are motivated to complete the tasks, but it also lent us popularity in the Haskell community and led to a historically high participation rate. Gamification has been shown to improve engagement and motivation. This has been harnessed by many research projects to improve participation in human studies \cite{He2014-vp}. We identify that studies on functional programming are often technical and intimidating; our use of gamification attracted historically high participation with a wide distribution of experience levels.

\subsection{Potential Applications}
\subsubsection{Programming Assistance}
We envision many ways GeckoGraph can be integrated into programming tools. GeckoGraph can visualize and inspect types in tooltips and pop-ups. It can be used to discover the mismatching parts of two conflicting types in type errors. It can be used to generate type expressions and edit type expressions structurally. In our post-study survey, the potential integration of text editors and programming assistance were the most requested use cases proposed by the participants.

\subsubsection{Documentation Assistance}
From what we have learned from our human study, GeckoGraph is well suited to support the documentation of the programming library and the API documentation. It works in tandem with the traditional text-based language and can be generated mechanically, making it possible to standardize with minimal effort. For documentation sites that allow searching by name (e.g., Hoogle \cite{Mitchell_undated-fb}), programmers often need to sieve through a list of identically named functions. For example, a simple Hoogle search for the name \texttt{make} shows a list of functions with vastly different usage and purpose. GeckoGraph can help speed up the selection process by providing a visual notation for each type, and programmers can use a visual grouping of colors, sizes, and positions to home in on the correct documentation page.

\subsubsection{Pedagogical Applications}

We believe that GeckoGraph can be a valuable tool in teaching techniques and theories in programming languages that are difficult to convey in plain language. In fact, many participants in our study reported that they had no prior knowledge of Haskell programming and that they could understand the programming concepts in the game and complete all the puzzles with the help of GeckoGraph.

Furthermore, the advanced features of GeckoGraph (Section~\ref{sec:benefits}) are also suitable for teaching and learning high-level functional programming concepts. Consider the \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension (Fig. \ref{fig:assoc}). Although the type signature is short, it is very difficult to trace the semantics mentally due to the number of variables, and their kinds are not obvious from the text-based notation. GeckoGraph makes understanding the type easier by exposing the ``hidden" higher-kinded types, revealing all the partially applied data types in play.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/assoc}
  \caption{\label{fig:assoc} The \texttt{assoc} function for day convolution \cite{Day1970-kb} in the Kan extension. Even for people who are not familiar with the exact definitions, it is easy to see that the variables \texttt{f},  \texttt{g}, and \texttt{h} are all high-kinded types, which is hard to tell from the text-based type annotation.}
\end{figure}

\section{Related work}
\subsection{Visualizing polymorphic types}
A similar half-enclosing structure was proposed in the visualization of types by Jung \cite{Jung2000-oc}. In Jung's notation, the type constructor half encloses its arguments, but the figure for the type constructor is placed on the bottom right (Fig. \ref{fig:jung}).  In contrast, GeckoGraph follows the natural reading order, allowing larger structures in a type signature to take precedence over smaller ones.

Compared to functions in Jung's notation,  GeckoGraph shows two major benefits. First, GeckoGraph naturally translates the general notion of a curried function. Partially, the application of a function can be read as removing the first one of its arguments. This is not the case with Jung's notation. Second, the shape of a function type remains consistent with the shape of normal data types. In Jung's design, a function \texttt{a -> b} looks sufficiently distinct from a data type \texttt{f a b}. This is important because, in functional languages, it is very common for abstraction to be drawn from function and normal data types. For example,  functions and lists both have a functor instance, and their inner values can be altered using a \texttt{fmap} function. The consistent shape of GeckoGraph makes this generalization easier to see visually.

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Jung}
  \caption{
        \label{fig:jung}
        Comparing the map function (\texttt{(a -> b) -> [a] -> [b]}) in text notation, GeckoGraph, and Jung's notation.
  }
\end{figure}

%Many studies \cite{Jun2000-ec, Jun2000-yu} explored the challenges of how human experts use polymorphic types. The authors then show a history-preserving type inference algorithm and explanation generator that can explain the types in human language. Similarly, \cite{Beaven1993-ay} showed a system that explains the infeasibility of types using human-like languages, explaining the type inference process using a series of "Why" questions and "How" questions. Both studies use natural language to reduce the challenge of understanding polymorphic types. While these studies are promising, explaining polymorphic types with natural language has limitations. They are generally verbose and have to use another visualization assist (text decoration, icons, or indentation) to clarify the naturally hierarchical information. A very similar graphical type representation \cite{Jung2000-oc} that utilizes blocks and colors was proposed. In the paper, the visualization uses 2D blocks to represent types and enclosures to represent type constructors and type arguments. We have discussed thoroughly the difference in design between Jung's notation and GeckoGraph. In addition, the evaluation of Jung's notation uses fixed questions and answers, while GeckoGraph is evaluated using real programming tasks.

\subsection{Visualization in programming}
Using visualization techniques to improve the comprehension of polymorphic types is not new. This is often practiced to represent document properties, runtime information, and static analysis results.  FluidEdt \cite{Ou2015-vr} displays heap graphs in the left margin. I3 \cite{Beck2015-my} offers search similarity and change history in compact block-based diagrams. Almeida et al. ~\cite{Almeida2022-bv} introduced a novel visualization to aid in understanding ownership and borrowing in the Rust language. While the field of graphical type representation is relatively narrow, it has been studied to some extent. Clerici et al. ~\cite{Clerici2013-ru} proposed a graph-based type inference system that shows the visual representations of unification states. GeckoGraph positions itself similarly to these projects, using color, shapes, symbols, and icons to provide easy-to-glance information. However, GeckoGraph focuses on type-level information, which no other research projects do. In addition, GeckoGraph is evaluated in a much larger study than the other projects, and GeckoGraph is evaluated with a wider range of experience levels.

\subsection{Visual vs Textual representation}

Many studies have compared the effectiveness of a visual-based programming environment with a textual-based one. Studies \cite{Noone2018-wl, Da_Silva_Ribeiro2014-tm, Cliburn2008-jo, Daly2011-is} found that compared to a purely textual programming language with similar positioning, students who were taught a visual programming language show greater confidence, better retention, and enjoyment in programming courses. While showing a similar trend, GeckoGraph experiments in the context of accompanying text-based notation rather than replacing it.

Many have studied the effect of visual augmentation, providing a visual representation of programming objects without removing the text-based notation. Greenfoot \cite{Montero2010-uh} allows visual and textual representations of programming concepts to be accessible to the learner. PILeT \cite{Alshaigy2015-wy}, providing a programming environment that is an adaptive presentation based on the user's preference. Both tools show positive results in the use of visual augmentation. Although similar to GeckoGraph in combining visual language and text-based programming environment, both studies evaluated the effect based on imperative languages (Java and Python), while our evaluation focused on the effect on a functional language (Haskell).

\section{Conclusion and Future Work}
In this paper, we propose GeckoGraph, a graphical notation for type annotations in functional programming languages. GeckoGraph aims to accompany traditional text-based type notation and to make reading, understanding, and comparing types easier. We conducted a large-scale human study using GeckoGraph compared to text-based type notation, the largest user study on functional programming we are aware of. The results of the study show that GeckoGraph helps improve programmers' ability to succeed in the programming tasks we designed, especially for novice programmers.

Our work in this area opens many new directions for future research.  In particular:

\noindent\textbf{In-the-wild Studies}
Although our experiment scenarios are drawn from real-world programming tasks, a certain level of variable control is still applied to remove the effect of familiarity with the tools and libraries. However, it is necessary to assess the usefulness of tools such as GeckoGraph by their real-life usage. To study this, different research methods should be used to study realistic usage and human experience. This may include field deployments or case studies.

\noindent\textbf{Alternative Type Visualization}
We strongly believe that visualization is an underutilized technique in this effort. GeckoGraph focuses on a faithful view of the tree structure of programming types. However, many more areas and types demand a human-focused approach. For instance, visualizing the ordinal relationship of subsumption or visualizing the numeric changes in dependently typed ``type programs".

\bibliographystyle{elsarticle-num} 
\bibliography{paperpile}

\appendix
\section{Game levels} \label{levels}
We provide all the level settings we used in our user study. The online game is still open source and available for evaluation \cite{Anonymous_undated-ne}. However, this can be attempted locally with a Haskell interpreter or even with a pen and paper. The target type is the desired type signature for the function \texttt{zeroToHero}. The available functions show a list of functions that are allowed to be used in the implementation. It is not required to use all the available functions, and it is not forbidden to use any other functions or variables outside the provided functions; even the Haskell prelude is not available.

\subsection{Level 1: Trial run}

\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f :: Zero a -> Hero a}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f z}
\end{itemize}

\subsection{Level 2: Assemble required}

\paragraph{Target Type} 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{runZero :: Zero a -> a}
    \item \texttt{mkHero :: a -> Hero a}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = mkHero (runZero z)}
\end{itemize}

\subsection{Level 3: Which path?}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a -> Hero (a, a)}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a -> Hero a}
    \item \texttt{f2 :: Zero a -> (a, a)}
    \item \texttt{f3 :: Hero a -> Hero (a, a)}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f1 \$ z}
\end{itemize}

\subsection{Level 4: A repeating pattern}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b -> Hero b b}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b -> Hero b a}
    \item \texttt{f2 :: Zero a a -> Hero a a}
    \item \texttt{f3 :: Zero a b -> Zero b a}
    \item \texttt{f4 :: Zero a b -> Zero b b}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f2 . f4 \$ z}
\end{itemize}

\subsection{Level 5: A perfect pair}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b -> Hero b b}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{fst :: (a, b) -> a}
    \item \texttt{snd :: (a, b) -> b}
    \item \texttt{f1 :: Zero a b -> Hero b a}
    \item \texttt{f2 :: Zero a a -> Hero a a}
    \item \texttt{f3 :: Zero a b -> Zero b a}
    \item \texttt{f4 :: Zero a b -> Zero b b}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = snd .f3 . f1 \$ z}
\end{itemize}

\subsection{Level 6: Monty Hall}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c -> Hero c a}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c-> Zero c b a}
    \item \texttt{f2 :: Zero a b c -> Zero a c c}
    \item \texttt{f3 :: Zero a b c -> Hero b c}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f1 . f2 \$ z}
\end{itemize}

\subsection{Level 7: TIE fighter}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c -> Hero c}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Hero (a -> b)}
    \item \texttt{f2 :: Zero a b c -> Hero (b -> c)}
    \item \texttt{f3 :: Zero a b c -> Hero a}
    \item \texttt{(<\$>) :: (a -> b) -> Hero a -> Hero b}
    \item \texttt{(<*>) :: Hero (a -> c) -> Hero a -> Hero c}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f2 z <*> (f1 z <*> f3 z)}
\end{itemize}

\subsection{Level 8: The middle man}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: (a -> d) -> (b -> d) -> (c -> d) -> Zero a b c ->  Hero a d c}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Zero c a b}
    \item \texttt{f2 :: Zero a b c -> Hero a b c}
    \item \texttt{fmap :: (c -> d) -> Zero a b c -> Zero a b d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero ad bd cd z = f2  . f1  . f1  . fmap bd  . f1 \$ z}
\end{itemize}

\subsection{Level 9: Split the difference}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero a b c d ->  Hero d d d d}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero a b c -> Zero c a b}
    \item \texttt{f2 :: Zero a b c -> Hero a b c}
    \item \texttt{fmap :: (c -> d) -> Zero a b c -> Zero a b d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero ad bd cd z = f2 \$ f1 \$ f1 \$ f3 \$ z}
\end{itemize}

\subsection{Level 10: The roller coaster}
\paragraph{Target Type } 
\begin{itemize}
    \item \texttt{zeroToHero :: Zero (a -> b -> c -> d) a b c  -> Hero d}
\end{itemize}

\paragraph{Available Functions} 
\begin{itemize}
    \item \texttt{f1 :: Zero (a -> b) a c d -> Zero () b c d}
    \item \texttt{f2 :: Zero a b c d -> Zero b c d a}
    \item \texttt{f3 :: Zero a b c d -> Hero d}
    \item \texttt{(\$) :: (a -> b) -> a -> b}
    \item \texttt{(.) :: (b -> c) -> (a -> b) -> a -> c}
\end{itemize}

\paragraph{Possible Solution} 
\begin{itemize}
    \item \texttt{zeroToHero z = f3 . f2 . f2 . f1 . f2 . f1 . f2 . f1 \$ z}
\end{itemize}

\end{document}
\endinput