### 2025-10-26 response.tex
Final updates

\documentclass{article}

\newcommand{\pointRaised}[2]{\medskip \noindent 
               \paragraph{#1} \textsl{#2}} 
\newcommand{\reply}{\noindent \textbf{Response}:\ }

\begin{document}

Dear Editor,

Thank you and the reviewers for the opportunity to revise our manuscript “GeckoGraph: A Visual Language for Polymorphic Types.”
We acknowledge the constructive and insightful feedback provided. We think they help improve the clarity, structure, and balance of our paper.
In this revised version, we have addressed all reviewer comments and made corresponding improvements throughout the manuscript. Below is a summary of major changes, followed by detailed point-by-point responses to each reviewer.

We hope our revisions satisfactorily address all concerns and that the paper is now suitable for publication in the Journal of Computer Languages.

Sincerely,

Authors of "GeckoGraph: A Visual Language for Polymorphic Types"
\section{Summary of Major Revisions}

\paragraph{Balanced claims} We have moderated the tone in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} to present a more balanced and evidence-based discussion, explicitly acknowledging both the strengths and limitations of our findings.

\paragraph{Terminology refinement} We have replaced references to the user study as a “game” with more accurate terms such as “gamified user study” or simply “user study,” to avoid overstating the level of interactivity.

\paragraph{New background section} We have added a new section, \textit{Type Systems in Functional Languages}, following the \textit{Introduction} to provide theoretical background and better motivate the design of GeckoGraph.

\paragraph{Section structure overviews} We have introduced short structural overviews at the beginning of the \textit{Related Work} and \textit{Discussion} sections to help guide readers through their organization.

\paragraph{Expanded related work} We have extended the \textit{Related Work} section to include additional references, notably the paper “Just TYPEical: Visualizing Common Function Type Signatures in R,” and used it to better position GeckoGraph among related visualization approaches.

\paragraph{Enhanced discussion} The \textit{Discussion} section has been revised to include more concrete examples demonstrating potential applications of GeckoGraph.

\paragraph{Concise figure captions} All figure captions have been shortened for clarity, readability, and stylistic consistency.

\paragraph{Improved flow between sections} A transitional paragraph has been added to connect the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3.1, improving the logical flow of ideas.

\paragraph{Clarified explanation of type classes} The explanation of type classes in Section~3.1 has been rewritten in simpler language with improved examples. Figure~3 now includes concrete illustrations to support the description.

\paragraph{Removed unused figure} We removed the previous Figure~4, as it did not contribute additional information to the paper.

\paragraph{Cosmetic and consistency edits} Grammar, spelling, LaTeX syntax, and stylistic consistency have been carefully reviewed and corrected throughout the manuscript.
\section{Detailed Responses to Reviewers}

\subsection*{Response to Reviewer 1}

\pointRaised{Comment 1:}{(Abstract, Introduction, Conclusion – Major) The manuscript’s tone in the Abstract and Conclusion appears overly strong given the presented evidence.}

\reply We appreciate this observation. The language in the \textit{Abstract}, \textit{Introduction}, and \textit{Conclusion} has been moderated to ensure that claims are proportionate to the results. The revised version now presents both the positive and negative aspects of our findings.

\pointRaised{Comment 2:}{(Overall) The term “game” used to describe the user study could be misleading.}

\reply We agree. The tasks in our study do not fully meet the criteria for a “game.” We now describe it more accurately as a “gamified user study” or simply a “user study,” emphasizing its educational and experimental nature.

\pointRaised{Comment 3:}{(Section 3.1) Clarify the explanation of type classes.}

\reply The subsection on type classes (Section~3.1) has been rewritten for clarity and supported by clearer examples. Figure~3 has also been improved to include illustrative examples that reinforce our description of GeckoGraph.

\pointRaised{Comment 4:}{(Section 3) Clarify the syntactic sugar of curried functions.}

\reply We now explicitly clarify that the function notation \texttt{a -> b -> c} desugars to \texttt{a -> (b -> c)}, which is consistently represented in GeckoGraph.

\pointRaised{Comment 5:}{(Section 3) Figure 3: The second higher-order function type {\tt ((a -> b) -> b)} is tagged as binary but is unary.}

\reply We agree that the previous figure was confusing. Figure~4 (previously Figure~3) has been simplified for readability.

\pointRaised{Comment 6:}{(Section 3) Issues with figures and references.}

\reply The previous Figure~4 has been removed. Incorrect cross-references such as “Fig.~2.2” have been corrected (now “Fig.~8”), and redundant phrases like “Appendix Appendix~A” have been fixed.

\pointRaised{Comment 7:}{(Section 5) Temper the narrative of qualitative results—the data show neutral user opinions.}

\reply We have clarified that participants’ feedback was overall neutral. Descriptors such as “slightly positive” and “moderately negative” are now used to reflect sentiment values more precisely.

\pointRaised{Comment 8:}{(Section 6) Clarify speculative claims in the Discussion.}

\reply We now explicitly mark speculative interpretations in Section~6 (\textit{Discussion}) using leading phrases such as “We hypothesize that...”.

\pointRaised{Comment 9:}{(Overall) Wording and grammar issues.}

\reply We have corrected numerous stylistic and grammatical issues, including:
“expressiveness power” → “expressiveness”,
“high-order function” → “higher-order function,”
and “GechoGraph” → “GeckoGraph.”  
Unclear expressions such as “tangibly” and “regular higher-order function” have been revised or replaced with precise terms, and unnecessary spaces before footnotes have been removed.

\subsection*{Response to Reviewer 2}

\pointRaised{Comment 1:}{(Section 1) Add a concluding paragraph in the Introduction describing the article’s organization.}

\reply We agree. A short paper-structure overview has been added at the end of Section~1 to guide the reader through the paper’s organization.

\pointRaised{Comment 2:}{(Section 2) The paper lacks sufficient background on type systems in functional languages.}

\reply We have added a new section titled \textit{Type Systems in Functional Languages} immediately after the \textit{Introduction}. This section introduces key theoretical concepts and motivates the design rationale for GeckoGraph.

\pointRaised{Comment 3:}{(Section 3) The paper’s flow between subsections feels abrupt.}

\reply We have added a transitional paragraph between the \textit{Design Goals} and \textit{Detailed GeckoGraph Features} subsections in Section~3 to improve narrative flow.

\pointRaised{Comment 4:}{(Section 4) Clarify user tasks and results in Figures 9 and 10.}

\reply We have clarified that the user in Figures~9 and~10 did not reach the correct solution and explained how GeckoGraph’s visual feedback can help identify and correct such mistakes. We also fixed a typo in Figure~9, changing “runHero z” to “runZero z.”

\pointRaised{Comment 5:}{(Section 4) Clarify whether participants indicate their experience once or before each run.}

\reply We clarified that participants report their programming experience once, before beginning the tasks.

\pointRaised{Comment 6:}{(Section 4) Add reference for “between-subject design.”}

\reply We now correctly cite Fisher, R.A. \textit{The Design of Experiments} for the phrase “between-subject design.”

\pointRaised{Comment 7:}{(Section 6) The Discussion section is abstract and lacks concrete examples.}

\reply We have revised the \textit{Discussion} section to include additional examples illustrating practical applications of GeckoGraph. Figure~14 has been added to support this discussion.

\pointRaised{Comment 8:}{(Section 6 and 7) Begin Sections~6 (“Discussion”) and~7 (“Related Work”) with overview paragraphs explaining purpose and structure.}

\reply Overview paragraphs have been added at the start of both sections to help orient readers and clarify their organization.

\pointRaised{Comment 9:}{(Section 7) The Related Work section could benefit from additional references and a clearer structure.}

\reply We have expanded and reorganized the \textit{Related Work} section, including new references such as “Just TYPEical: Visualizing Common Function Type Signatures in R.” A short overview paragraph now outlines the section’s focus and structure.

\pointRaised{Comment 10:}{(Section 6) Distinguish between empirical findings and interpretations in the Discussion.}

\reply We now explicitly distinguish between observed results and speculative interpretations in the \textit{Discussion}, using phrases such as “We hypothesize that...” to signal interpretive commentary.

\pointRaised{Comment 11:}{(Overall) Usage of figures—some captions too long, Figure~4 unclear.}

\reply Figure captions throughout the paper have been shortened for conciseness. The previous Figure~4, which did not add new insight, has been removed.

\pointRaised{Comment 12:}{(Overall) Minor wording and grammar issues.}

\reply The manuscript has been thoroughly proofread for grammar, formatting, and coherence.  
Examples include corrected figure references (e.g., “Fig.~2.2” → “Fig.~8”), removal of redundant text (“Appendix Appendix~A” → “Appendix~A”), and clarification of unclear terms such as “approximity” and “during type errors.”

\end{document}